---
editor_options: 
  markdown: 
    wrap: 72
---

![logo_ironhack_blue
7](https://user-images.githubusercontent.com/23629340/40541063-a07a0a8a-601a-11e8-91b5-2f13e4e6b441.png)

# Lab \| Classes, functions & operators in R

This lab is designed to help you practice creating custom classes,
defining operators, and writing functions in R. Follow the instructions
and complete the tasks below.

## Task 1: Create a Custom S3 Class

1.  Define a custom S3 class called "Circle" that represents a circle.
    The class should store the radius of the circle.

2.  Write a constructor function Circle() that takes the radius as input
    and assigns the class "Circle" to the object.

3.  Write a print.Circle() method to display the radius and area of the
    circle when the object is printed. Use the formula for the area of a
    circle: area = π \* radius\^2.

    ```{r}
    Circle <- function(radius, x, y){
      if ((!is.numeric(radius)) || (!is.numeric(x)) || (!is.numeric(y))) {
        stop("Please, input numeric value.")
      }
      structure(list(radius = radius, x = x, y = y), class = "Circle")
    } 

    print.Circle <- function(obj) {
      cat("Radius = ", obj$radius, "\n")
      area <- 3.1415926 * obj$radius * obj$radius
      cat("Area = ", area)
    }

    circumference <- function(obj) {
      UseMethod("circumference")
    }

    circumference.Circle <- function(obj){
      circumfc <- 3.1415926 * 2 * obj$radius
      print(circumfc)
    }

    my_circle <- Circle(radius=5, x=9, y=0)
    print(my_circle)
    ```

```{r}
my_non_Circle <- Circle(radius="Five", x=0, y=0)
print(my_non_Circle)
```

**Example:**

``` r
my_circle <- Circle(5)
print(my_circle)
```

**Output:** <br> Type: "Circle object" <br> Radius: 5<br> Area:
78.53982<br>

## Task 2: Define a Custom Operator

1.  Define a custom operator %\>% that calculates the distance between
    the centers of two circles and checks if they intersect. The
    operator should return TRUE if the circles intersect and FALSE
    otherwise. The formula for the distance between two circles is:
    distance = sqrt((x2 - x1)\^2 + (y2 - y1)\^2).

    -   Assume that the circles are centered at (x1, y1) and (x2, y2),
        respectively.

    -   Two circles intersect if the distance between their centers is
        less than or equal to the sum of their radii.

    ```{r}
        `%>%` <- function(circle1, circle2){
          if (!inherits(circle1, "Circle") || !inherits(circle2, "Circle")){
            stop("Both operands must be of class 'Circle'.")
          }
          distance <- sqrt((circle1$x - circle2$x)^2 + (circle1$y - circle2$y)^2)
          if (distance > circle1$radius + circle2$radius) {
            return(FALSE)
          } else {
            return(TRUE)
          }
        }
    ```

2.  Modify the Circle class to include x and y coordinates for the
    center of the circle.

    **Example:**
    `r  circle1 <- Circle(radius = 3, x = 0, y = 0)  circle2 <- Circle(radius = 4, x = 5, y = 0)  circle1 %> circle2`
    **Output:** TRUE

    ```{r}
    circle1 <- Circle(radius = 3, x = 0, y = 0) 
    circle2 <- Circle(radius = 4, x = 5, y = 0)
    interf <- circle1 %>% circle2
    print(interf)
    ```

```{r}
    circle1 <- Circle(radius = 1, x = 0, y = 0) 
    circle2 <- Circle(radius = 1, x = 5, y = 0)
    interf <- circle1 %>% circle2
    print(interf)
```

## Task 3: Write a Function for the Class

1.  Write a function circumference() that calculates the circumference
    of a Circle object. The formula for the circumference is:
    circumference = 2 \* π \* radius.

2.  Add this function as a method for the Circle class so that it can be
    called directly on objects of the class.

    **Example:**
    `r  my_circle <- Circle(radius = 5)  circumference(my_circle)`
    **Output:** 31.41593

    ```{r}
    circumference(my_circle)
    ```

## Task 4: Create a Custom S4 Class

1.  Define a custom S4 class called "Rectangle" that represents a
    rectangle. The class should store the length and width of the
    rectangle.

2.  Write a constructor function Rectangle() that takes the length and
    width as input and creates an object of the "Rectangle" class.

3.  Write a show() method for the Rectangle class to display the length,
    width, and area of the rectangle when the object is printed.

    ```{r}

    setClass("Rectangle",
      slots = list(
        length = "numeric",
        width = "numeric"
      )
    )


    Rectangle <- function(length, width) {
      if (!is.numeric(length) || !is.numeric(width) || length <= 0 || width <= 0) {
        stop("Length and width must be positive numeric values.")
      }
      new("Rectangle", length = length, width = width)
    }


    setMethod("show", "Rectangle", function(object) {
        area <- object@length * object@width
        cat("Rectangle:\n")
        cat("Length:", object@length, "\n")
        cat("Width:", object@width, "\n")
        cat("Area:", area, "\n")
      }
    )


    rect <- Rectangle(5, 10)
    print(rect)
    ```

    **Example:**
    `r  my_rectangle <- Rectangle(length = 4, width = 6)  show(my_rectangle)`
    **Output:** <br> Type: "Rectangle object" <br> Length: 4 <br> Width:
    6 <br> Area: 24<br>

## Task 5: Define a Custom Operator for the S4 Class

1.  Define a custom operator %==% that compares two Rectangle objects
    and checks if they have the same area.

2.  The operator should return TRUE if the areas are equal and FALSE
    otherwise.

    **Example:**
    `r  rectangle1 <- Rectangle(length = 4, width = 6)  rectangle2 <- Rectangle(length = 3, width = 8)  rectangle1 %==% rectangle2`
    **Output:** TRUE

    ```{r}
    "%==%" <- function(rec1, rec2) {
      if (!inherits(rec1, "Rectangle") || (!inherits(rec2, "Rectangle"))) {
        stop("Both operands must be of class 'Rectangle'.")
      }
      if (rec1@length * rec1@width == rec2@length * rec2@width) {
            return(TRUE)
          } else {
            return(FALSE)
          }
    }

    rectangle1 <- Rectangle(length = 4, width = 6)  
    rectangle2 <- Rectangle(length = 3, width = 8)  
    rectangle1 %==% rectangle2

    rectangle3 <- Rectangle(length = 2, width = 5)  
    rectangle4 <- Rectangle(length = 3, width = 10)  
    rectangle3 %==% rectangle4
    ```

### Resources

-   [R Documentation on S3 Classes](https://adv-r.hadley.nz/s3.html)
-   [R Documentation on S4 Classes](https://adv-r.hadley.nz/s4.html)
-   [R
    Operators](https://stat.ethz.ch/R-manual/R-devel/library/base/html/Arithmetic.html)

## Deliverables

-   Submitted notebook (or file) with your responses to each of the
    exercises.

## Submission

-   Upon completion, add your deliverables to git.
-   Then commit git and push your branch to the remote.
-   Make a pull request and paste the PR link in the submission field in
    the Student Portal.

<br>

**Good luck!**
